---
title: "The Box Office"
execute: 
  echo: false
  warning: false
format:
  html: default
  # pdf: default
---

## Admissions
Despite months of closures at the start of the year and subsequent restrictions in the form of capacity limits and rules on social distancing, cinemas enjoyed significantly greater audience numbers in 2021 than in 2020.

Admissions for the year totalled 74 million, up 68% compared with 2020, but still well below the level of attendances seen in 2019 ([@fig-admissions-html]{.content-visible when-format="html"}[@fig-admission-pdf]{.content-visible when-format="pdf"}).

```{r}
# load packages

library(tidyverse)

# import data and skip comment lines

admissions <- read_csv(
  file = "data/admissions.csv", 
  comment = "#"
  )
```

::: {.content-visible when-format="html"}
```{r}
#| label: fig-admissions-html
#| fig-cap: Annual UK cinema admissions, 2012-2021.
#| column: page

library(plotly)

fig1 <- ggplot(
  mapping = aes(
    x = Year,
    y = `Admissions (million)`,
  )
) +
  geom_point(data = admissions, size = 2, alpha = 1, color = "grey") +
  geom_line(data = admissions, size = 1, alpha = 0.5, color = "purple") +
  expand_limits(y = 0) +
  labs(
    title = "Number of admissions (million)",
    x = "Year",
    y = "",
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(
      face = "bold",
      margin = margin(b = 10)
    ),
    plot.title.position = "plot",
    plot.margin = margin(15, 10, 10, 15)
  )
ggplotly(fig1) %>%
  layout(hovermode = "x")
```
:::

::: {.content-visible when-format="pdf"}
```{r}
#| label: fig-admissions-pdf
#| fig-cap: Annual UK cinema admissions, 2012-2021.

fig1 <- ggplot(
  data = admissions,
  mapping = aes(
    x = Year,
    y = `Admissions (million)`,
  )
) +
  geom_point(size = 2, alpha = 1, color = "grey") +
  geom_text(
    aes(label = `Admissions (million)`), nudge_y = -10
  ) +
  geom_line(size = 1, alpha = 0.5, color = "purple") +
  expand_limits(y = 0) +
  labs(
    title = "Number of admissions (million)",
    x = "Year",
    y = "",
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(
      face = "bold",
      margin = margin(b = 10)
    ),
    plot.title.position = "plot",
    plot.margin = margin(15, 10, 10, 15)
  )
fig1
```
:::

[@tbl-markets-html]{.content-visible when-format="html"}[@tbl-markets-pdf]{.content-visible when-format="pdf"} shows admissions data for the largest world and European film markets between 2019 and 2021. As with the UK, the majority of territories saw increased attendances in 2021 in comparison with 2020 but cinema-going in every territory remained well below the level of admissions in 2019. The two territories reporting numbers of ticket sales in 2021 closest to those of 2019 were China (down only 32%) and Russia (-34%) while those with the greatest shortfall between 2019 and 2021 were Indonesia (-82%), Italy (-74%) and South Korea (-73%).

According to Omdia, global cinema admissions in 2021 totalled 3.3 billion, an increase from 1.9 billion in 2020, but still 57% down on the 7.6 billion recorded in 2019.

```{r}
library(knitr)

markets <- read_csv(
  file = "data/largest-markets.csv", 
  comment = "#"
  )
```

::: {.content-visible when-format="html"}
```{r}
#| label: tbl-markets-html
#| tbl-cap:  Largest film markets by admissions, 2019-2021 (ranked by 2021 admissions).

library(reactable)

reactable(markets, defaultPageSize = 15)
```
:::

::: {.content-visible when-format="pdf"}
```{r}
#| label: tbl-markets-pdf
#| tbl-cap:  Largest film markets by admissions, 2019-2021 (ranked by 2021 admissions).

kable(markets, format = "latex")
```
:::